---
title: "WM 2026 — Live Model Benchmark"
subtitle: "Out-of-sample evaluation & simulated P&L · Auto-updated via GitHub Actions"
date: today
format:
  dashboard:
    theme: darkly
    nav-buttons:
      - icon: github
        href: https://github.com/vincent-rgb-cpu/WM_2026
---

```{r setup, include=FALSE}
library(dplyr)
library(plotly)
library(DT)
library(scales)

# Paths relative to project root (dashboard/ is rendered from root in CI)
ROOT <- if (file.exists("output/wc2026_match_log.csv")) "." else ".."

match_log <- tryCatch(
  readr::read_csv(file.path(ROOT, "output/wc2026_match_log.csv"),
                  show_col_types = FALSE),
  error = function(e) NULL
)
pnl <- tryCatch(
  readr::read_csv(file.path(ROOT, "output/financial_benchmark.csv"),
                  show_col_types = FALSE),
  error = function(e) NULL
)
tourney <- tryCatch(
  readr::read_csv(file.path(ROOT, "output/tournament_probabilities.csv"),
                  show_col_types = FALSE),
  error = function(e) NULL
)
metrics <- tryCatch(
  readr::read_csv(file.path(ROOT, "output/evaluation_metrics.csv"),
                  show_col_types = FALSE),
  error = function(e) NULL
)
fixture_preds <- tryCatch(
  readr::read_csv(file.path(ROOT, "output/fixture_predictions.csv"),
                  show_col_types = FALSE),
  error = function(e) NULL
)

has_matches <- !is.null(match_log) && nrow(match_log) > 0
has_pnl     <- !is.null(pnl)       && nrow(pnl) > 1

# Ensure odds columns exist (match_log written before odds integration won't have them)
if (has_matches) {
  if (!"bookmaker" %in% names(match_log)) match_log$bookmaker <- NA_character_
  if (!"real_odds"  %in% names(match_log)) match_log$real_odds  <- NA_real_
  if (!"edge"       %in% names(match_log)) match_log$edge        <- NA_real_
}

# Identify the next unplayed fixture
next_match <- NULL
if (!is.null(fixture_preds) && nrow(fixture_preds) > 0) {
  played_ids <- if (has_matches) match_log$match_id else integer(0)
  upcoming   <- fixture_preds %>%
    filter(!(match_id %in% played_ids)) %>%
    arrange(date) %>%
    slice(1)
  if (nrow(upcoming) > 0) next_match <- upcoming
}

TEAM_FLAGS <- c(
  "Algeria"                = "\U0001F1E9\U0001F1FF",
  "Argentina"              = "\U0001F1E6\U0001F1F7",
  "Australia"              = "\U0001F1E6\U0001F1FA",
  "Austria"                = "\U0001F1E6\U0001F1F9",
  "Belgium"                = "\U0001F1E7\U0001F1EA",
  "Bosnia and Herzegovina" = "\U0001F1E7\U0001F1E6",
  "Brazil"                 = "\U0001F1E7\U0001F1F7",
  "Canada"                 = "\U0001F1E8\U0001F1E6",
  "Cape Verde"             = "\U0001F1E8\U0001F1FB",
  "Colombia"               = "\U0001F1E8\U0001F1F4",
  "Croatia"                = "\U0001F1ED\U0001F1F7",
  "Curaçao"           = "\U0001F1E8\U0001F1FC",
  "Czech Republic"         = "\U0001F1E8\U0001F1FF",
  "DR Congo"               = "\U0001F1E8\U0001F1E9",
  "Ecuador"                = "\U0001F1EA\U0001F1E8",
  "Egypt"                  = "\U0001F1EA\U0001F1EC",
  "England"                = "\U0001F1EC\U0001F1E7",
  "France"                 = "\U0001F1EB\U0001F1F7",
  "Germany"                = "\U0001F1E9\U0001F1EA",
  "Ghana"                  = "\U0001F1EC\U0001F1ED",
  "Haiti"                  = "\U0001F1ED\U0001F1F9",
  "Iran"                   = "\U0001F1EE\U0001F1F7",
  "Iraq"                   = "\U0001F1EE\U0001F1F6",
  "Ivory Coast"            = "\U0001F1E8\U0001F1EE",
  "Japan"                  = "\U0001F1EF\U0001F1F5",
  "Jordan"                 = "\U0001F1EF\U0001F1F4",
  "Mexico"                 = "\U0001F1F2\U0001F1FD",
  "Morocco"                = "\U0001F1F2\U0001F1E6",
  "Netherlands"            = "\U0001F1F3\U0001F1F1",
  "New Zealand"            = "\U0001F1F3\U0001F1FF",
  "Norway"                 = "\U0001F1F3\U0001F1F4",
  "Panama"                 = "\U0001F1F5\U0001F1E6",
  "Paraguay"               = "\U0001F1F5\U0001F1FE",
  "Portugal"               = "\U0001F1F5\U0001F1F9",
  "Qatar"                  = "\U0001F1F6\U0001F1E6",
  "Saudi Arabia"           = "\U0001F1F8\U0001F1E6",
  "Scotland"               = "\U0001F1EC\U0001F1E7",
  "Senegal"                = "\U0001F1F8\U0001F1F3",
  "South Africa"           = "\U0001F1FF\U0001F1E6",
  "South Korea"            = "\U0001F1F0\U0001F1F7",
  "Spain"                  = "\U0001F1EA\U0001F1F8",
  "Sweden"                 = "\U0001F1F8\U0001F1EA",
  "Switzerland"            = "\U0001F1E8\U0001F1ED",
  "Tunisia"                = "\U0001F1F9\U0001F1F3",
  "Turkey"                 = "\U0001F1F9\U0001F1F7",
  "United States"          = "\U0001F1FA\U0001F1F8",
  "Uruguay"                = "\U0001F1FA\U0001F1FE",
  "Uzbekistan"             = "\U0001F1FA\U0001F1FF"
)

team_label <- function(name) {
  flag <- TEAM_FLAGS[name]
  if (!is.na(flag)) paste0(flag, " ", name) else name
}
```

# Predictions

## Row

### Upcoming Fixtures & Model Predictions

```{r}
if (!is.null(fixture_preds) && nrow(fixture_preds) > 0) {
  played_ids   <- if (has_matches) match_log$match_id else integer(0)
  upcoming_all <- fixture_preds %>%
    filter(!(match_id %in% played_ids)) %>%
    arrange(date)

  if (nrow(upcoming_all) > 0) {
    pred_df <- upcoming_all %>%
      rowwise() %>%
      mutate(
        Date  = format(as.Date(date), "%a %d %b"),
        Stage = if (stage == "group") paste0("Grp ", group)
                else tools::toTitleCase(gsub("_", " ", stage)),
        Home  = team_label(home_team),
        Away  = team_label(away_team),
        Pick  = dplyr::case_when(
          pred_result == "home_win" ~ team_label(home_team),
          pred_result == "draw"     ~ "Draw",
          TRUE                      ~ team_label(away_team)
        ),
        `P(H)` = round(p_home_win * 100, 1),
        `P(D)` = round(p_draw     * 100, 1),
        `P(A)` = round(p_away_win * 100, 1)
      ) %>%
      ungroup() %>%
      select(Date, Stage, Home, Away, `P(H)`, `P(D)`, `P(A)`, Pick)

    DT::datatable(
      pred_df,
      escape   = FALSE,
      rownames = FALSE,
      options  = list(
        pageLength = 30,
        dom        = "tip",
        scrollX    = TRUE,
        columnDefs = list(
          list(className = "dt-center", targets = c(0, 1, 4, 5, 6)),
          list(width = "22%", targets = c(2, 3)),
          list(width = "7%",  targets = c(4, 5, 6))
        )
      ),
      class = "table-dark table-sm table-hover"
    ) %>%
      DT::formatStyle(
        "P(H)",
        background         = DT::styleColorBar(c(0, 100), "#4a6fa5"),
        backgroundSize     = "98% 55%",
        backgroundRepeat   = "no-repeat",
        backgroundPosition = "center"
      ) %>%
      DT::formatStyle(
        "P(D)",
        background         = DT::styleColorBar(c(0, 100), "#5a6268"),
        backgroundSize     = "98% 55%",
        backgroundRepeat   = "no-repeat",
        backgroundPosition = "center"
      ) %>%
      DT::formatStyle(
        "P(A)",
        background         = DT::styleColorBar(c(0, 100), "#c0392b"),
        backgroundSize     = "98% 55%",
        backgroundRepeat   = "no-repeat",
        backgroundPosition = "center"
      ) %>%
      DT::formatStyle("Pick", fontWeight = "bold", color = "#28a745") %>%
      DT::formatStyle(c("P(H)", "P(D)", "P(A)"), color = "#e0e0e0")
  } else {
    cat("All fixtures have been played — check the Financial Simulation tab for results.")
  }
} else {
  cat("Fixture predictions not available.")
}
```

# ML Metrics

## Row {height=22%}

### Next Match — Model Prediction

```{r}
if (!is.null(next_match)) {
  home  <- next_match$home_team
  away  <- next_match$away_team
  ph    <- round(next_match$p_home_win * 100, 1)
  pd    <- round(next_match$p_draw     * 100, 1)
  pa    <- round(next_match$p_away_win * 100, 1)
  pred  <- next_match$pred_result

  stage_label <- if (next_match$stage == "group")
    paste0("Group ", next_match$group) else tools::toTitleCase(gsub("_", " ", next_match$stage))
  date_label  <- format(as.Date(next_match$date), "%a %d %b %Y")

  bar_cols <- c(
    ifelse(pred == "home_win", "#28a745", "#4a6fa5"),
    ifelse(pred == "draw",     "#28a745", "#6c757d"),
    ifelse(pred == "away_win", "#28a745", "#c0392b")
  )

  # Factor levels control display order (home on top, away on bottom)
  labels <- factor(c(home, "Draw", away), levels = c(away, "Draw", home))

  plot_ly() %>%
    add_bars(
      x    = c(ph, pd, pa),
      y    = labels,
      orientation  = "h",
      marker       = list(color = bar_cols),
      text         = paste0(c(ph, pd, pa), "%"),
      textposition = "inside",
      insidetextanchor = "middle",
      textfont     = list(size = 14, color = "#fff")
    ) %>%
    layout(
      title = list(
        text = paste0(home, " vs ", away,
                      "  ·  ", date_label,
                      "  ·  ", stage_label),
        font = list(size = 14, color = "#adb5bd"),
        x    = 0,
        xanchor = "left"
      ),
      xaxis = list(range = c(0, 100), showgrid = FALSE, zeroline = FALSE,
                   showticklabels = FALSE, title = ""),
      yaxis = list(title = "", tickfont = list(size = 13)),
      showlegend    = FALSE,
      paper_bgcolor = "rgba(0,0,0,0)",
      plot_bgcolor  = "rgba(0,0,0,0)",
      font   = list(color = "#fff"),
      margin = list(l = 10, r = 60, t = 45, b = 10)
    )
} else {
  cat("No upcoming fixtures found.")
}
```

## Row {height=18%}

```{r}
#| content: valuebox
#| title: "WC 2026 Accuracy"
acc_live <- if (has_matches) percent(mean(match_log$correct), 0.1) else "—"
list(icon = "bullseye", color = "success", value = acc_live)
```

```{r}
#| content: valuebox
#| title: "WC 2026 Log-Loss"
ll_live <- if (has_matches) round(mean(match_log$log_loss), 3) else "—"
list(icon = "activity", color = "info", value = ll_live)
```

```{r}
#| content: valuebox
#| title: "Held-out Accuracy (2021+)"
acc_hist <- if (!is.null(metrics))
  percent(metrics$accuracy[metrics$model == "xgboost"], 0.1) else "—"
list(icon = "graph-up", color = "primary", value = acc_hist)
```

```{r}
#| content: valuebox
#| title: "Matches Evaluated"
n_matches <- if (has_matches) nrow(match_log) else 0
list(icon = "calendar-check", color = "secondary", value = n_matches)
```

## Row {height=40%}

### Calibration — Predicted vs Actual {width=60%}

```{r}
if (has_matches) {
  # Bin predicted probabilities and compare to empirical frequencies
  cal <- match_log %>%
    tidyr::pivot_longer(c(p_home_win, p_draw, p_away_win),
                        names_to = "outcome", values_to = "pred_prob") %>%
    mutate(
      outcome_label = recode(outcome,
        p_home_win = "Home Win", p_draw = "Draw", p_away_win = "Away Win"),
      bin = cut(pred_prob, breaks = seq(0, 1, 0.1), include.lowest = TRUE)
    ) %>%
    group_by(outcome_label, bin) %>%
    summarise(
      mean_pred = mean(pred_prob),
      mean_act  = mean(actual_result == recode(outcome_label,
        "Home Win" = "home_win", "Draw" = "draw", "Away Win" = "away_win")),
      n = n(), .groups = "drop"
    ) %>%
    filter(n >= 2)

  plot_ly(cal, x = ~mean_pred, y = ~mean_act, color = ~outcome_label,
          size = ~n, type = "scatter", mode = "markers",
          marker = list(opacity = 0.8)) %>%
    add_lines(x = c(0,1), y = c(0,1), line = list(dash = "dash", color = "grey"),
              showlegend = FALSE, inherit = FALSE) %>%
    layout(
      xaxis = list(title = "Predicted Probability", range = c(0,1)),
      yaxis = list(title = "Empirical Frequency",   range = c(0,1)),
      legend = list(orientation = "h"),
      paper_bgcolor = "rgba(0,0,0,0)", plot_bgcolor = "rgba(0,0,0,0)",
      font = list(color = "#fff")
    )
} else {
  cat("No completed WC-2026 matches yet.")
}
```

### Model vs Baselines {width=40%}

```{r}
if (!is.null(metrics)) {
  metrics_long <- metrics %>%
    tidyr::pivot_longer(c(accuracy, log_loss, brier),
                        names_to = "metric", values_to = "value") %>%
    mutate(metric = recode(metric,
      accuracy = "Accuracy ↑", log_loss = "Log-Loss ↓", brier = "Brier ↓"))

  plot_ly(metrics_long, x = ~model, y = ~value, color = ~model,
          type = "bar", facet_col = ~metric) %>%
    layout(
      barmode = "group",
      xaxis = list(title = ""),
      yaxis = list(title = "Score"),
      showlegend = FALSE,
      paper_bgcolor = "rgba(0,0,0,0)", plot_bgcolor = "rgba(0,0,0,0)",
      font = list(color = "#fff")
    )
} else {
  cat("Metrics file not found.")
}
```

## Row {height=40%}

### Match-by-Match Log-Loss {width=60%}

```{r}
if (has_matches) {
  ml_sorted <- match_log %>%
    arrange(date) %>%
    mutate(
      label    = paste(home_team, "vs", away_team),
      cum_ll   = cummean(log_loss),
      match_no = row_number()
    )

  plot_ly() %>%
    add_bars(data = ml_sorted, x = ~match_no, y = ~log_loss,
             name = "Per-Match Log-Loss", marker = list(color = "#6c757d")) %>%
    add_lines(data = ml_sorted, x = ~match_no, y = ~cum_ll,
              name = "Cumulative Mean", line = list(color = "#17a2b8", width = 2)) %>%
    layout(
      xaxis = list(title = "Match Number"),
      yaxis = list(title = "Log-Loss"),
      legend = list(orientation = "h"),
      paper_bgcolor = "rgba(0,0,0,0)", plot_bgcolor = "rgba(0,0,0,0)",
      font = list(color = "#fff")
    )
} else {
  cat("No completed matches yet.")
}
```

### Tournament Odds (Top 12) {width=40%}

```{r}
if (!is.null(tourney)) {
  top12 <- tourney %>%
    arrange(desc(Win_World_Cup)) %>%
    head(12) %>%
    mutate(Team = forcats::fct_reorder(Team, Win_World_Cup))

  plot_ly(top12, x = ~Win_World_Cup, y = ~Team, type = "bar",
          orientation = "h",
          marker = list(color = ~Win_World_Cup, colorscale = "Viridis")) %>%
    layout(
      xaxis = list(title = "P(Win Cup) %"),
      yaxis = list(title = ""),
      paper_bgcolor = "rgba(0,0,0,0)", plot_bgcolor = "rgba(0,0,0,0)",
      font = list(color = "#fff")
    )
} else {
  cat("Tournament probabilities not found.")
}
```

# Financial Simulation

> **Methodology:** A bet is placed on the model's highest-probability outcome when
> (1) **real pre-match market odds** are available from The Odds API (preferred source: Pinnacle)
> and (2) the model's **edge** — `model_prob − 1 / real_odds` — exceeds **3 %**.
> Stake = Quarter-Kelly of current bankroll, capped at 10 %.
> Matches without real odds are logged but excluded from the P&L simulation.

## Row {height=20%}

```{r}
#| content: valuebox
#| title: "Starting Bankroll"
list(icon = "bank", color = "secondary", value = "1,000 units")
```

```{r}
#| content: valuebox
#| title: "Current Bankroll"
bk_now <- if (has_pnl) comma(round(tail(pnl$bankroll, 1))) else "—"
list(icon = "cash-stack", color = "success", value = bk_now)
```

```{r}
#| content: valuebox
#| title: "Simulated ROI"
roi <- if (has_pnl) {
  r <- (tail(pnl$bankroll, 1) - 1000) / 1000
  paste0(if (r >= 0) "+" else "", percent(r, 0.1))
} else "—"
col <- if (!is.null(match_log) && has_pnl &&
           tail(pnl$bankroll, 1) >= 1000) "success" else "danger"
list(icon = "percent", color = col, value = roi)
```

```{r}
#| content: valuebox
#| title: "Bets Placed"
n_bets <- if (has_pnl) tail(pnl$n_bets, 1) else "—"
list(icon = "lightning-charge", color = "info", value = n_bets)
```

## Row {height=45%}

### Cumulative Bankroll {width=65%}

```{r}
if (has_pnl) {
  plot_ly(pnl, x = ~date, y = ~bankroll, type = "scatter", mode = "lines+markers",
          line = list(color = "#28a745", width = 2),
          marker = list(color = "#28a745", size = 6),
          fill = "tozeroy", fillcolor = "rgba(40,167,69,0.15)") %>%
    add_lines(x = range(pnl$date), y = c(1000, 1000),
              line = list(dash = "dash", color = "grey", width = 1),
              showlegend = FALSE, inherit = FALSE) %>%
    layout(
      xaxis = list(title = "Date"),
      yaxis = list(title = "Bankroll (units)"),
      paper_bgcolor = "rgba(0,0,0,0)", plot_bgcolor = "rgba(0,0,0,0)",
      font = list(color = "#fff")
    )
} else {
  cat("No P&L data yet.")
}
```

### Return per Bet {width=35%}

```{r}
if (has_matches) {
  bet_data <- match_log %>%
    filter(kelly_stake > 0) %>%
    arrange(date) %>%
    mutate(
      label  = paste(home_team, "vs", away_team),
      colour = if_else(bet_return > 0, "Win", "Loss")
    )

  plot_ly(bet_data, x = ~seq_along(bet_return), y = ~bet_return,
          type = "bar", color = ~colour,
          colors = c("Win" = "#28a745", "Loss" = "#dc3545"),
          text = ~label, hoverinfo = "text+y") %>%
    layout(
      xaxis = list(title = "Bet #"),
      yaxis = list(title = "Return (units)"),
      showlegend = TRUE,
      legend = list(orientation = "h"),
      paper_bgcolor = "rgba(0,0,0,0)", plot_bgcolor = "rgba(0,0,0,0)",
      font = list(color = "#fff")
    )
} else {
  cat("No bets placed yet.")
}
```

## Row {height=35%}

### Full Match Log

```{r}
if (has_matches) {
  match_log %>%
    arrange(date) %>%
    mutate(
      Score       = paste0(home_score, "–", away_score),
      Predicted   = pred_result,
      Actual      = actual_result,
      `P(actual)` = round(p_actual, 3),
      `Log-Loss`  = round(log_loss, 3),
      Correct     = if_else(correct, "✓", "✗"),
      Bookmaker   = ifelse(is.na(bookmaker), "—", bookmaker),
      Odds        = ifelse(is.na(real_odds), "—", as.character(round(real_odds, 2))),
      Edge        = ifelse(is.na(edge), "—", paste0(round(edge * 100, 1), "%")),
      Stake       = round(kelly_stake, 1),
      Return      = round(bet_return, 1),
      Bankroll    = round(bankroll, 1)
    ) %>%
    select(Date = date, Home = home_team, Away = away_team, Score,
           Predicted, Actual, `P(actual)`, `Log-Loss`, Correct,
           Bookmaker, Odds, Edge, Stake, Return, Bankroll) %>%
    DT::datatable(
      options = list(pageLength = 16, dom = "tip", scrollX = TRUE),
      rownames = FALSE,
      class    = "table-dark table-sm"
    ) %>%
    DT::formatStyle("Correct",
      color = DT::styleEqual(c("✓", "✗"), c("#28a745", "#dc3545")))
}
```
