---
title: "WM 2026 — Live Model Benchmark"
subtitle: "Out-of-sample evaluation & simulated P&L · Auto-updated via GitHub Actions"
date: today
format:
  dashboard:
    theme: darkly
    nav-buttons:
      - icon: github
        href: https://github.com/vincent-rgb-cpu/WM_2026
---

```{r setup, include=FALSE}
library(dplyr)
library(plotly)
library(DT)
library(scales)

# Paths relative to project root (dashboard/ is rendered from root in CI)
ROOT <- if (file.exists("output/wc2026_match_log.csv")) "." else ".."

match_log <- tryCatch(
  readr::read_csv(file.path(ROOT, "output/wc2026_match_log.csv"),
                  show_col_types = FALSE),
  error = function(e) NULL
)
pnl <- tryCatch(
  readr::read_csv(file.path(ROOT, "output/financial_benchmark.csv"),
                  show_col_types = FALSE),
  error = function(e) NULL
)
tourney <- tryCatch(
  readr::read_csv(file.path(ROOT, "output/tournament_probabilities.csv"),
                  show_col_types = FALSE),
  error = function(e) NULL
)
metrics <- tryCatch(
  readr::read_csv(file.path(ROOT, "output/evaluation_metrics.csv"),
                  show_col_types = FALSE),
  error = function(e) NULL
)

has_matches <- !is.null(match_log) && nrow(match_log) > 0
has_pnl     <- !is.null(pnl)       && nrow(pnl) > 1
```

# ML Metrics

## Row {height=20%}

```{r}
#| content: valuebox
#| title: "WC 2026 Accuracy"
acc_live <- if (has_matches) percent(mean(match_log$correct), 0.1) else "—"
list(icon = "bullseye", color = "success", value = acc_live)
```

```{r}
#| content: valuebox
#| title: "WC 2026 Log-Loss"
ll_live <- if (has_matches) round(mean(match_log$log_loss), 3) else "—"
list(icon = "activity", color = "info", value = ll_live)
```

```{r}
#| content: valuebox
#| title: "Held-out Accuracy (2021+)"
acc_hist <- if (!is.null(metrics))
  percent(metrics$accuracy[metrics$model == "xgboost"], 0.1) else "—"
list(icon = "graph-up", color = "primary", value = acc_hist)
```

```{r}
#| content: valuebox
#| title: "Matches Evaluated"
n_matches <- if (has_matches) nrow(match_log) else 0
list(icon = "calendar-check", color = "secondary", value = n_matches)
```

## Row {height=40%}

### Calibration — Predicted vs Actual {width=60%}

```{r}
if (has_matches) {
  # Bin predicted probabilities and compare to empirical frequencies
  cal <- match_log %>%
    tidyr::pivot_longer(c(p_home_win, p_draw, p_away_win),
                        names_to = "outcome", values_to = "pred_prob") %>%
    mutate(
      outcome_label = recode(outcome,
        p_home_win = "Home Win", p_draw = "Draw", p_away_win = "Away Win"),
      bin = cut(pred_prob, breaks = seq(0, 1, 0.1), include.lowest = TRUE)
    ) %>%
    group_by(outcome_label, bin) %>%
    summarise(
      mean_pred = mean(pred_prob),
      mean_act  = mean(actual_result == recode(outcome_label,
        "Home Win" = "home_win", "Draw" = "draw", "Away Win" = "away_win")),
      n = n(), .groups = "drop"
    ) %>%
    filter(n >= 2)

  plot_ly(cal, x = ~mean_pred, y = ~mean_act, color = ~outcome_label,
          size = ~n, type = "scatter", mode = "markers",
          marker = list(opacity = 0.8)) %>%
    add_lines(x = c(0,1), y = c(0,1), line = list(dash = "dash", color = "grey"),
              showlegend = FALSE, inherit = FALSE) %>%
    layout(
      xaxis = list(title = "Predicted Probability", range = c(0,1)),
      yaxis = list(title = "Empirical Frequency",   range = c(0,1)),
      legend = list(orientation = "h"),
      paper_bgcolor = "rgba(0,0,0,0)", plot_bgcolor = "rgba(0,0,0,0)",
      font = list(color = "#fff")
    )
} else {
  cat("No completed WC-2026 matches yet.")
}
```

### Model vs Baselines {width=40%}

```{r}
if (!is.null(metrics)) {
  metrics_long <- metrics %>%
    tidyr::pivot_longer(c(accuracy, log_loss, brier),
                        names_to = "metric", values_to = "value") %>%
    mutate(metric = recode(metric,
      accuracy = "Accuracy ↑", log_loss = "Log-Loss ↓", brier = "Brier ↓"))

  plot_ly(metrics_long, x = ~model, y = ~value, color = ~model,
          type = "bar", facet_col = ~metric) %>%
    layout(
      barmode = "group",
      xaxis = list(title = ""),
      yaxis = list(title = "Score"),
      showlegend = FALSE,
      paper_bgcolor = "rgba(0,0,0,0)", plot_bgcolor = "rgba(0,0,0,0)",
      font = list(color = "#fff")
    )
} else {
  cat("Metrics file not found.")
}
```

## Row {height=40%}

### Match-by-Match Log-Loss {width=60%}

```{r}
if (has_matches) {
  ml_sorted <- match_log %>%
    arrange(date) %>%
    mutate(
      label    = paste(home_team, "vs", away_team),
      cum_ll   = cummean(log_loss),
      match_no = row_number()
    )

  plot_ly() %>%
    add_bars(data = ml_sorted, x = ~match_no, y = ~log_loss,
             name = "Per-Match Log-Loss", marker = list(color = "#6c757d")) %>%
    add_lines(data = ml_sorted, x = ~match_no, y = ~cum_ll,
              name = "Cumulative Mean", line = list(color = "#17a2b8", width = 2)) %>%
    layout(
      xaxis = list(title = "Match Number"),
      yaxis = list(title = "Log-Loss"),
      legend = list(orientation = "h"),
      paper_bgcolor = "rgba(0,0,0,0)", plot_bgcolor = "rgba(0,0,0,0)",
      font = list(color = "#fff")
    )
} else {
  cat("No completed matches yet.")
}
```

### Tournament Odds (Top 12) {width=40%}

```{r}
if (!is.null(tourney)) {
  top12 <- tourney %>%
    arrange(desc(Win_World_Cup)) %>%
    head(12) %>%
    mutate(Team = forcats::fct_reorder(Team, Win_World_Cup))

  plot_ly(top12, x = ~Win_World_Cup, y = ~Team, type = "bar",
          orientation = "h",
          marker = list(color = ~Win_World_Cup, colorscale = "Viridis")) %>%
    layout(
      xaxis = list(title = "P(Win Cup) %"),
      yaxis = list(title = ""),
      paper_bgcolor = "rgba(0,0,0,0)", plot_bgcolor = "rgba(0,0,0,0)",
      font = list(color = "#fff")
    )
} else {
  cat("Tournament probabilities not found.")
}
```

# Financial Simulation

> **Methodology:** A bet is placed on the model's highest-probability outcome when
> (1) **real pre-match market odds** are available from The Odds API (preferred source: Pinnacle)
> and (2) the model's **edge** — `model_prob − 1 / real_odds` — exceeds **3 %**.
> Stake = Quarter-Kelly of current bankroll, capped at 10 %.
> Matches without real odds are logged but excluded from the P&L simulation.

## Row {height=20%}

```{r}
#| content: valuebox
#| title: "Starting Bankroll"
list(icon = "bank", color = "secondary", value = "1,000 units")
```

```{r}
#| content: valuebox
#| title: "Current Bankroll"
bk_now <- if (has_pnl) comma(round(tail(pnl$bankroll, 1))) else "—"
list(icon = "cash-stack", color = "success", value = bk_now)
```

```{r}
#| content: valuebox
#| title: "Simulated ROI"
roi <- if (has_pnl) {
  r <- (tail(pnl$bankroll, 1) - 1000) / 1000
  paste0(if (r >= 0) "+" else "", percent(r, 0.1))
} else "—"
col <- if (!is.null(match_log) && has_pnl &&
           tail(pnl$bankroll, 1) >= 1000) "success" else "danger"
list(icon = "percent", color = col, value = roi)
```

```{r}
#| content: valuebox
#| title: "Bets Placed"
n_bets <- if (has_pnl) tail(pnl$n_bets, 1) else "—"
list(icon = "lightning-charge", color = "info", value = n_bets)
```

## Row {height=45%}

### Cumulative Bankroll {width=65%}

```{r}
if (has_pnl) {
  plot_ly(pnl, x = ~date, y = ~bankroll, type = "scatter", mode = "lines+markers",
          line = list(color = "#28a745", width = 2),
          marker = list(color = "#28a745", size = 6),
          fill = "tozeroy", fillcolor = "rgba(40,167,69,0.15)") %>%
    add_lines(x = range(pnl$date), y = c(1000, 1000),
              line = list(dash = "dash", color = "grey", width = 1),
              showlegend = FALSE, inherit = FALSE) %>%
    layout(
      xaxis = list(title = "Date"),
      yaxis = list(title = "Bankroll (units)"),
      paper_bgcolor = "rgba(0,0,0,0)", plot_bgcolor = "rgba(0,0,0,0)",
      font = list(color = "#fff")
    )
} else {
  cat("No P&L data yet.")
}
```

### Return per Bet {width=35%}

```{r}
if (has_matches) {
  bet_data <- match_log %>%
    filter(kelly_stake > 0) %>%
    arrange(date) %>%
    mutate(
      label  = paste(home_team, "vs", away_team),
      colour = if_else(bet_return > 0, "Win", "Loss")
    )

  plot_ly(bet_data, x = ~seq_along(bet_return), y = ~bet_return,
          type = "bar", color = ~colour,
          colors = c("Win" = "#28a745", "Loss" = "#dc3545"),
          text = ~label, hoverinfo = "text+y") %>%
    layout(
      xaxis = list(title = "Bet #"),
      yaxis = list(title = "Return (units)"),
      showlegend = TRUE,
      legend = list(orientation = "h"),
      paper_bgcolor = "rgba(0,0,0,0)", plot_bgcolor = "rgba(0,0,0,0)",
      font = list(color = "#fff")
    )
} else {
  cat("No bets placed yet.")
}
```

## Row {height=35%}

### Full Match Log

```{r}
if (has_matches) {
  match_log %>%
    arrange(date) %>%
    mutate(
      Score       = paste0(home_score, "–", away_score),
      Predicted   = pred_result,
      Actual      = actual_result,
      `P(actual)` = round(p_actual, 3),
      `Log-Loss`  = round(log_loss, 3),
      Correct     = if_else(correct, "✓", "✗"),
      Bookmaker   = if_else(is.na(bookmaker), "—", bookmaker),
      Odds        = if_else(is.na(real_odds), "—", as.character(round(real_odds, 2))),
      Edge        = if_else(is.na(edge), "—", percent(edge, 0.1)),
      Stake       = round(kelly_stake, 1),
      Return      = round(bet_return, 1),
      Bankroll    = round(bankroll, 1)
    ) %>%
    select(Date = date, Home = home_team, Away = away_team, Score,
           Predicted, Actual, `P(actual)`, `Log-Loss`, Correct,
           Bookmaker, Odds, Edge, Stake, Return, Bankroll) %>%
    DT::datatable(
      options = list(pageLength = 16, dom = "tip", scrollX = TRUE),
      rownames = FALSE,
      class    = "table-dark table-sm"
    ) %>%
    DT::formatStyle("Correct",
      color = DT::styleEqual(c("✓", "✗"), c("#28a745", "#dc3545")))
}
```
